Ae. albopictus CPP manuscript

### 2019-10-25 MS.Rmd
Maps and methods

---
title: "Ae albopictus CPP manuscript"
header-includes: #allows you to add in your own Latex packages
- \usepackage{float} #use the 'float' package
- \floatplacement{figure}{H} #make every figure with caption = h
- \usepackage{caption}
- \captionsetup[figure]{labelformat=empty}
- \captionsetup[table]{labelformat=empty}
output:
  pdf_document: default
  html_document: default
  word_document: default
---

```{r eval=FALSE, message=FALSE, include=FALSE}
install.packages("tidyverse") #1.2.1
install.packages("drc") #3.0.1
install.packages("stringr")
install.packages("geosphere")
install.packages("lubridate")
install.packages(c("maps", "mapdata"))
devtools::install_github("dkahle/ggmap")
install.packages("scatterpie")
install.packages("ggedit")
install.packages("cowplot")
```

```{r message=FALSE, echo=FALSE, include=FALSE}
library("drc") #3.0.1
library("knitr") #1.25
library("tidyverse") #1.2.1
library("stringr") #1.4.0
library("geosphere") #1.5.10
library("lubridate") #1.7.4
library("maps")
library("mapdata")
library("scatterpie")
library("ggedit")
library("cowplot")
```

@@

us_map<-ggplot() + geom_polygon(data = usa, aes(x=long, y = lat, group=group), fill=NA, color="blue") + 
  coord_fixed(1.3) + geom_point(data = filter(pops, country=="US"), aes(x = longitude, y = latitude), size = 1) + xlim(c(-90, -65)) +
  geom_label(data=filter(pops, country=="US"), aes(x=longitude, y=latitude, label=pop), hjust=0, label.padding=unit(0.1, "lines")) +
  theme_bw()

japan_map<-ggplot() + geom_polygon(data = jp, aes(x=long, y = lat, group=group), fill=NA, color="red") + 
  coord_fixed(1.3) + geom_point(data = filter(pops, country=="JP", only_2008=="n"), aes(x = longitude, y = latitude), size = 1)  +
  geom_label(data=filter(pops, country=="JP"), aes(x=longitude, y=latitude, label=pop), hjust=0, label.padding=unit(0.1, "lines")) +
  theme_bw() + theme(axis.title.x=element_blank(), axis.title.y=element_blank())

@@

DI_pies_2018<-plot_grid(remove_geom(us_map, "label") + geom_scatterpie(data=filter(DI_info, country=="US", year=="2018"), aes(x=longitude, y=latitude, r=0.5), cols=c("Diapause", "Nondiapause")) + scale_fill_manual(values=c("black","white")) + theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank(), axis.title.y=element_blank(), legend.position="none", plot.margin = unit(c(0, 0, 0, 0), "cm")), remove_geom(japan_map, "label") + geom_scatterpie(data=filter(DI_info, country=="JP", year=="2018"), aes(x=longitude, y=latitude, r=0.5), cols=c("Diapause", "Nondiapause")) + scale_fill_manual(values=c("black","white")) + theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank(), axis.title.y=element_blank(), legend.position="none", plot.margin = unit(c(0, 0, 0, 0), "cm")), ncol=1, align="hv")
```

@@

- Diapause is an important overwintering adaptation in many organisms
- This adaptation comes with trade-offs, but exactly what they are is unclear (winter phenotypes of diapausing organisms; life history strategies)
- Climate change may have unpredictable effects on growing seasons
- Invasive species provide natural experiments for rapid adaptation
- Aedes albopictus worldwide spread: diapause played a role 
- What is the continuing role of diapause in light of environmental change, continuing introductions, competing selective pressures? 
- We recapitulated 2008 CPP study

@@

### Mosquito population sampling and rearing
Mosquito populations were sampled in the eastern USA and Japan in 2018, at latitudes ranging from `r min(pops$latitude)` to `r max(pops$latitude)`. Sampling locations are described in Table 1. Locations in the US were largely identical to locations sampled in 2008. [More details about field sampling.] At least 50 male and 50 female mosquitoes were used to initiate laboratory populations, except for SEN, which had X females and Y males.

@@

Populations were maintained in the laboratory for 3 or 4 generations prior to the experiment at 21C and 80% humidity under a 16:8h light:dark cycle. Larvae were reared in distilled water and fed a slurry of dog food and brine shrimp, while adults were fed organic raisins and blood-fed by human arm. Initial laboratory generations of larvae were reared in plastic storage containers, with one population per container and densities varying by population. Pupae were then transferred into one adult cage per population. Populations generally had at least 50 female and 50 male adults per generation, but the JAC population experienced visibly high mortality during the F2 generation. The larval generation immediately prior to the experiment was reared in 15 cm plastic petri dishes containing 30-40 larvae each, X dishes per population. Additional details of mosquito husbandry can be found in Supplementary File 1.

@@

Due to the large number of populations and photoperiods, the experiment was conducted in two blocks. Each block contained half of the populations from each country; populations across the full range of latitudes were interspersed between the blocks. At beginning of each block, eggs were hatched by submerging oviposition papers in water with larval food for two days. Larvae were distributed among X 15 cm petri dishes containing 90 ml of a mixture of distilled water and larval food (5 L of water to 10 mL of food slurry) at a density of 30-40 larvae per dish. Larvae were transferred to clean dishes containing fresh water/food mixture three times a week until pupation. Larvae were kept in the controlled temperature room at 21 degrees, 80 percent humidity and 16:8 hours light:dark until they produced pupae.

Pupae from each population were collected, pooled across dishes, and haphazardly distributed amongst thirteen 1.5 L adult cages, corresponding to 12 photoperiod cabinets and an additional short-day (8h light) incubator. The total number of pupae added to each cage ranged from X to Y; subsequent pupal and adult mortality was not monitored. Approximately one week after the last pupae were added, cages were blood-fed, and a dark oviposition cup half-filled with water and lined with unbleached paper towel was added to each cage. Oviposition papers were collected 5-7 days after blood-feeding and replaced every 2-3 days; a total of four oviposition papers were collected from each cage. Several cages were fed again and an additional oviposition paper was collected; see Supplementary File 1 for details. Oviposition papers were dried gently and stored in the photoperiod cabinets for 10 to 21 days until the diapause scoring procedure.

@@

```{r table2, echo=FALSE}
kable(anova(dimodel), digits=3, caption="Table 2")
```

```{r fig2, fig.cap="Figure 2: Diapause incidence (2018) by country. Black sections of pie charts represent diapausing fraction; white fractions represent non-diapausing. Blue points represent US diapause incidence; red points represent Japan.", echo=FALSE, message=FALSE}
plot_grid(DI_pies_2018, country_DI_2018, ncol=2, labels="AUTO")
```

CPP values correlated strongly with latitude in the US and Japan. Latitudinal clines were nearly parallel to those observed in 2008 (Figure 3), as verified by the lack of a significant interaction term between year and latitude (Table 3).

@@

However, the intercepts of the latitudinal clines significantly differed between both the US and Japan, and between 2008 and 2018 (Figure 4). Latitude-specific CPP was higher in Japan than the US, and higher in 2018 compared to 2008. This implies earlier entry into diapause at the same latitude in Japan than in the US and earlier entry into diapause in the same locations in 2018 than in 2008.

>>>>

---
title: "Ae. albopictus CPP manuscript"
header-includes: #allows you to add in your own Latex packages
- \usepackage{float} #use the 'float' package
- \floatplacement{figure}{H} #make every figure with caption = h
- \usepackage{caption}
- \captionsetup[figure]{labelformat=empty}
- \captionsetup[table]{labelformat=empty}
output:
  word_document: default
  html_document: default
  pdf_document: default
---

```{r eval=FALSE, message=FALSE, include=FALSE}
install.packages("tidyverse") #1.2.1
install.packages("drc") #3.0.1
install.packages("stringr")
install.packages("geosphere")
install.packages("lubridate")
install.packages(c("maps", "mapdata"))
devtools::install_github("dkahle/ggmap")
install.packages("scatterpie")
install.packages("ggedit")
install.packages("ggrepel")
install.packages("cowplot")
```

```{r message=FALSE, echo=FALSE, include=FALSE}
library("drc") #3.0.1
library("knitr") #1.25
library("tidyverse") #1.2.1
library("stringr") #1.4.0
library("geosphere") #1.5.10
library("lubridate") #1.7.4
library("maps")
library("mapdata")
library("scatterpie")
library("ggedit")
library("ggrepel")
library("cowplot")
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(dpi=300)
```

@@

us_map<-ggplot() + geom_polygon(data = usa, aes(x=long, y = lat, group=group), fill=NA, color="blue") + 
  coord_fixed(1.3) + geom_point(data = filter(pops, country=="US"), aes(x = longitude, y = latitude), size = 1) + xlim(c(-85, -65)) +
  geom_label(data=filter(pops, country=="US"), aes(x=longitude, y=latitude, label=pop), hjust=0, label.padding=unit(0.1, "lines")) +
  theme_bw()

japan_map<-ggplot() + geom_polygon(data = jp, aes(x=long, y = lat, group=group), fill=NA, color="red") + 
  coord_fixed(1.3) + geom_point(data = filter(pops, country=="JP", only_2008=="n"), aes(x = longitude, y = latitude), size = 1)  + xlim(c(125, 147)) +
  geom_label_repel(data=filter(pops, country=="JP", only_2008=="n"), aes(x=longitude, y=latitude, label=pop), hjust=0, label.padding=unit(0.1, "lines")) +
  theme_bw() + theme(axis.title.x=element_blank(), axis.title.y=element_blank())

@@

DI_pies_2018<-plot_grid(remove_geom(us_map, "label") %>% remove_geom("point") + ylim(24,48) + geom_scatterpie(data=filter(DI_info, country=="US", year=="2018"), aes(x=longitude, y=latitude, r=.7, alpha=0.95), cols=c("Diapause", "Nondiapause")) + scale_fill_manual(values=c("black","white")) + theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank(), axis.title.y=element_blank(), legend.position="none"),

remove_geom(japan_map, "label") %>% remove_geom("point") + ylim(24,48) + geom_scatterpie(data=filter(DI_info, country=="JP", year=="2018"), aes(x=longitude, y=latitude, r=.7, alpha=0.95), cols=c("Diapause", "Nondiapause")) + scale_fill_manual(values=c("black","white")) + theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank(), axis.title.y=element_blank(), legend.position="none"), ncol=2, align="hv")
```

@@

- Diapause is an important overwintering adaptation in many organisms
- This adaptation comes with trade-offs, but exactly what they are is not well specified (discuss winter phenotypes of diapausing organisms; life history strategies)
- Climate change may have unpredictable effects on growing seasons
- Invasive species provide natural experiments for rapid adaptation
- Aedes albopictus worldwide spread: diapause played a role 
- What is the continuing role of diapause in light of environmental change, continuing introductions, competing selective pressures? 
- We recapitulated 2008 CPP study

@@

### Mosquito population sampling and rearing
Mosquito populations were sampled in the eastern USA and Japan in 2018, at latitudes ranging from `r min(pops$latitude)` to `r max(pops$latitude)`. Sampling locations are described in Table 1. Locations in the US were largely identical to locations sampled in 2008. [More details about field sampling.] At least 50 male and 50 female mosquitoes were used to initiate laboratory populations, except for SEN, which had 37 females and 39 males initially.

@@

Populations were maintained in the laboratory for 3 or 4 generations prior to the experiment at 21C and 80% humidity under a 16:8h light:dark cycle. Larvae were reared in distilled water and fed a slurry of dog food and brine shrimp, while adults were fed organic raisins and blood-fed by human arm. Initial laboratory generations of larvae were reared in plastic storage containers, with one population per container and densities varying by population. Pupae were then transferred into one adult cage per population. Populations generally had at least 50 female and 50 male adults per generation, but the JAC population experienced visibly high mortality during the F2 generation. The larval generation immediately prior to the experiment was reared in 15 cm plastic petri dishes containing 30-40 larvae each, 10 dishes per population. Additional details of mosquito husbandry can be found in Supplementary File 1.

@@

Due to the large number of populations and photoperiods, the experiment was conducted in two blocks. Each block contained half of the populations from each country; populations across the full range of latitudes were interspersed between the blocks. At beginning of each block, eggs were hatched by submerging oviposition papers in water with larval food for two days. Larvae were distributed among 40 15 cm petri dishes containing 90 ml of a mixture of distilled water and larval food (5 L of water to 10 mL of food slurry) at a density of 30-40 larvae per dish. Larvae were transferred to clean dishes containing fresh water/food mixture three times a week until pupation. Larvae were kept in the controlled temperature room at 21 degrees, 80 percent humidity and 16:8 hours light:dark until they produced pupae.

Pupae from each population were collected, pooled across dishes, and split amongst thirteen 1.5 L adult cages, corresponding to 12 photoperiod cabinets and an additional short-day (8h light) incubator. The total number of pupae added to each cage ranged from 52 to 106; pupae were not sexed and subsequent pupal and adult mortality was not monitored. Approximately one week after the last pupae were added, cages were blood-fed, and a dark oviposition cup half-filled with water and lined with unbleached paper towel was added to each cage. Oviposition papers were collected 5-7 days after blood-feeding and replaced every 2-3 days; a total of four oviposition papers were collected from each cage. Several cages were fed again and an additional oviposition paper was collected; see Supplementary File 1 for details. Oviposition papers were dried gently and stored in the photoperiod cabinets for 10 to 21 days until the diapause scoring procedure.

@@

```{r table2, echo=FALSE}
kable(anova(dimodel), digits=3, caption="Table 2: Effects of latitude, country and year (2008 vs 2018) on diapause incidence.")
```

```{r fig2A, fig.cap="Figure 2A: Diapause incidence (2018) by country. Dark sections of pie charts represent diapausing fraction; white fractions represent non-diapausing.", echo=FALSE, message=FALSE}
DI_pies_2018
```

```{r fig.cap="Figure 2B: Diapause incidence (2018) by country. Blue points represent US diapause incidence; red points represent Japan.", echo=FALSE, message=FALSE}
country_DI_2018
```

CPP values correlated strongly with latitude in the US and Japan. Latitudinal clines were nearly parallel to those observed in 2008 (Figure 3), as verified by the lack of a significant interaction term between year and latitude (Table 3). However, the intercepts of the latitudinal clines significantly differed between both the US and Japan, and between 2008 and 2018 (Figure 4). Latitude-specific CPP was higher in Japan than the US, and higher in 2018 compared to 2008. This implies earlier entry into diapause at the same latitude in Japan than in the US and earlier entry into diapause in the same locations in 2018 than in 2008.

@@
